---
title: "Exploratory Structural Equation Modelling"
author: "Jay Jeffries"
date: "1/10/2023"
output: 
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
pkgdown:
  as_is: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load("lavaan", "devtools", "dataxray", "tidyverse", "skimr", "Hmisc", "psych")
# devtools::install_github("agstn/dataxray")

# Name of first dataset "Tests"
# Renaming columns
Tests <- Tests |> rename(
  verbal1 = x1,
  verbal2 = x2,
  verbal3 = x3,
  math1 = y1,
  math2 = y2,
  math3 = y3,
  )

Tests$verbal <- rowSums(Tests[,1:3], na.rm = T)/3
Tests$math <- rowSums(Tests[,4:6], na.rm = T)/3

# Name of second dataset = "HolzingerSwineford1939"

HolzingerSwineford1939 <- HolzingerSwineford1939 |> rename(
  visual1 = x1,
  visual2 = x2,
  visual3 = x3,
  text1 = x4,
  text2 = x5,
  text3 = x6,
  speed1 = x7,
  speed2 = x8,
  speed3 = x9
  )

HolzingerSwineford1939$visual <- rowSums(HolzingerSwineford1939[,7:9])/3
HolzingerSwineford1939$text <- rowSums(HolzingerSwineford1939[,10:12])/3
HolzingerSwineford1939$speed <- rowSums(HolzingerSwineford1939[,13:15])/3
```

## Data Exploration

Using `{skimr}` package for data exploration.

#### Dataset 1 Descriptives
```{r}
skim(Tests)
```

#### Dataset 2 Descriptives
```{r}
skim(HolzingerSwineford1939)
```

## Bivariate Correlation

Using `{Hmisc}` package for correlation matrix.

* Found via average or summing scores 
* Simple model structure (items map only to single construct)
* Measured without measurement error
* Substantially inflates size of correlation

#### Dataset 1 Correlation Matrix
```{r}
corr_mat1 <- rcorr(as.matrix(Tests))
corr_mat1
```

#### Dataset 2 Correlation Matrix
```{r}
corr_mat2 <- rcorr(as.matrix(HolzingerSwineford1939[, 7:18]))
corr_mat2
```

## Confirmatory Factor Analysis Approach

Using the `{lavaan}` package.

* Items load only on *one* factor -- the one they are hypothesized to load onto
  * Restrictive, independent clustering of items
* Cross-loadings constrained to zero
* Moderately inflates size of correlation

#### Dataset 1 Model
```{r}
model1 <- '
Verbal =~ verbal1 + verbal2 + verbal3
Math =~ math1 + math2 + math3

Verbal ~~ Math
'

fit_model1 <- cfa(model1, data = Tests)
summary(fit_model1, standardized = T, fit.measures = T)
```

#### Dataset 2 Model
```{r}
model12.1<- '
Visual =~ visual1 + visual2 + visual3
Text =~ text1 + text2 + text3

Visual ~~ Text
'

fit_model2.1 <- cfa(model12.1, data = HolzingerSwineford1939)
summary(fit_model2.1, standardized = T, fit.measures = T)
```

## Exploratory Structural Equation Modelling Approach

[Site for syntax](https://sites.psu.edu/yuweili/research-methods/bifactor-esem/)
[Other site for syntax](https://msilvestrin.me/post/esem/)

* Uninflated factor correlation
* Cross-loadings estimated, even if small
* Improvement to fit indices

#### EFA for Model 1 Start Values
```{r}
model1_EFA <- fa(Tests[,1:6], nfactors = 2, rotate = "geominQ", fm = 'ML', delta = .5)
model1_EFA$loadings
```

#### Model 1 ESEM
```{r}
model1_2 <- '
Verbal =~ verbal1 + verbal2 + verbal3 + math1 + math2 + math3
Math =~ math1 + math2 + math3 + verbal1 + verbal2 + verbal3

Verbal ~~ Math
'

fit_model1_2 <- cfa(model1_2, data = Tests)
summary(fit_model1_2, standardized = T, fit.measures = T)
```

#### EFA for Model 2 Start Values
```{r}
model2.1_EFA <- fa(HolzingerSwineford1939[,7:12], nfactors = 3, rotate = "geominQ", fm = 'ML', delta = .5)
model2.1_EFA$loadings
```

```{r}
fit <- efa(data = HolzingerSwineford1939[,7:15], nfactors = 1:3)
summary(fit)
```


#### Model 2 ESEM
```{r}
model1_2.1 <- '
Visual =~ start(0.564)*visual1 + start(0.448)*visual2 + start(0.762)*visual3 + 
          start(0.102)*text1 + start()*text2 + start(0.145)*text3

Text =~ start(0.219)*text1 + text2 + text3 + 
        start(0.816)*visual1 + start(0.881)*visual2 + start(0.788)*visual3

Visual ~~ Text
'

fit_model1_2.1 <- cfa(model1_2.1, data = HolzingerSwineford1939, std.lv = T)
summary(fit_model1_2.1, standardized = T, fit.measures = T)
```

```{r}
model <- '
    # efa block 1
    efa("efa1")*f1 + 
    efa("efa1")*f2 =~ x1 + x2 + x3 + x4 + x5 + x6

    # efa block 2
    efa("efa2")*f3 + 
    efa("efa2")*f4 =~ y1 + y2 + y3 + y4 + y5 + y6

    # cfa block
    f5 =~ z7 + z8 + z9
    f6 =~ z10 + z11 + z12

    # regressions
    f3 ~ f1 + f2
    f4 ~ f3
'

fit.model <- cfa(model, data = HolzingerSwineford1939, std.lv = T)
summary(fit.model, standardized = T, fit.measures = T)
```