---
title: "Packages Of Interest"
author: "Jay Jeffries"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Colorblind-Friendly Palettes
### {col4all}

```{r cars, echo = F}
library(pacman)
p_load("cols4all", "colorspace", "abind", "shiny", "purrr", "tidyverse", "here", "readxl", "ggplot2", install = T)

# c4a_gui()
```

## Northern Color Palette
### [{nord}]("https://cran.r-project.org/web/packages/nord/readme/README.html")

```{r, fig.align = "center"}
library(nord)

par(mfrow=c(8, 2), lheight = 2, mar=rep(1, 4), adj = 0)

walk(names(nord_palettes), nord_show_palette)
```

## Interactive ggplot2 Data Visualizations
### {esquisse} - French for *"sketch"*

```{r pressure, echo=FALSE}
library(esquisse)

# squisse::esquisser()
```

## Download Data Files via Markdown Button
### [{downloadthis}]("https://fmmattioni.github.io/downloadthis/articles/downloadthis.html")

```{r}
library(downloadthis)

mtcars %>% # can also download a list of datasets via list(mtcars, iris) %>% but this can only save as .xlsx or .Rds file types
  download_this(
    output_name = "mtcars dataset",
    output_extension = ".xlsx", # can substitute .csv or 
    button_label = "Download mtcars Data as xlsx",
    button_type = "primary", # can use "default", "success", "info", "warning", "danger" for different button colors 
    has_icon = TRUE,
    icon = "fa fa-save"
  )

download_link(
  link = "https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/jjeffries13/Technostress/blob/main/Technostress_Codebook.csv",
  button_label = "Download csv Codebook File",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Simplified Exploratory Data Analysis (EDA)
### [{explore}]("https://github.com/rolkra/explore")

```{r}
library(explore)
library(tibble)

explore(sleep)

# sleep |> report(output_dir = "/Users/jayjeffries/Desktop/R Resources", target = extra, split = F)

explore(iris)
```

## Interactive Data Visualizations
### [{loon}]("https://great-northern-diver.github.io/loon/")

```{r}
devtools::install_github("great-northern-diver/loon", subdir = "R")
library(loon)

h <- l_hist(quakes$depth,
            yshows = "density",
            showBinHandle = TRUE,
            xlabel = "depth",
            title = "Tonga trench earthquakes",
            linkingGroup = "quakes")

p <- l_plot(x = quakes$long, y = quakes$lat,
            xlabel = "longitude", ylabel = "latitude",
            linkingGroup = "quakes",
            title = "Tonga trench earthquakes")

with(l_scale3D(quakes),
     l_plot3D(long, lat, depth, linkingGroup = "quakes")
)

NZFijiMap <- maps::map("world2", regions = c("New Zealand", "Fiji"), plot = FALSE)
l_layer(p, NZFijiMap,
        label = "New Zealand and Fiji",
        color = "forestgreen",
        index = "end")
# facet `p`
levels <- rep(NA, nrow(quakes))
levels[quakes$mag < 5 & quakes$mag >= 4] <- "Light"
levels[quakes$mag < 6 & quakes$mag >= 5] <- "Moderate"
levels[quakes$mag >= 6] <- "Strong"
l_facet(p, by = levels, 
        linkingGroup = "quakes")

```

## Correlation Network
### [Psychometric Summaries]("https://eiko-fried.com/creating-basic-psychometric-summaries-in-r/")

```{r}
p_load("psych", "qgraph", "bootnet", "OpenMx", "EGAnet", "lavaan", "sna", install = T)

data <- read_excel(path = here("./Likert Data/data/TeacherBurnout copy.xlsx")) 
data <- data |>
  select(-districtID, -schoolID)

cor_mat <- cor(data)
# vechs() gives you the strict half-vectorization of an input matrix as a column vector; i.e., the correlations above the diagonal

corr_desc <- summary(vechs(cor_mat)) # average correlation is 0.386
as.numeric(corr_desc[4])

qgraph(cor_mat, cut = 0, layout = "spring", 
       title = paste("Correlation matrix, mean correlation = ", round(as.numeric(corr_desc[4]), digits = 2), sep = " "),
       edge.width = .75, posCol = "steelblue4", negCol = "firebrick2")
# optional argument for colorblind theme; theme = "colorblind"

# Exploratory Graph Analysis (EGA) estimates the number of dimensions of a given dataset or correlation matrix based 
# Edges correspond to partial correlation coefficients between two variables after conditioning on all other variables in the network
# If items derive from same latent variable, two indicators cannot become independent after conditioning on observed variables because they are both caused by a latent variable (i.e., the edge strength between two indicators should not be zero).
network_ega <- EGA(data, 
                   corr = "pearson",
                   model = "glasso", # graphical least absolute shrinkage and selection operator (glasso); computes partial correlation matrix
                   algorithm = "walktrap") # walktrap algorithm is used to find the number of clusters of the partial correlation matrix
summary(network_ega)
# Golino et al. (2020) showed that the EGA method performs as well as the best factor-analytic techniques

boot_EGA <- bootEGA(data, iter= 1000,  type = "resampling")

boot_EGA$summary.table
boot_EGA$frequency

itemStability(boot_EGA)
# All items elicit stability within their "item community"
# Of the 1000 replications, TE5 remained in the same item community 93% of the time
```

## Plot Annotations
### Using [{ggannotate}]("https://github.com/MattCowgill/ggannotate")

```{r}
# remotes::install_github("mattcowgill/ggannotate")
library(ggannotate)

plot <- mtcars |>
  ggplot(aes(x = cyl)) +
  geom_bar()

# the below function opens a Shiny app window to add plot annotations via point-and-click methods
# annotation options include text, labels (text in a textbox), rectangles, and curved arrows
ggannotate(plot)

p <- ggplot(mtcars, 
            aes(x = wt, y = mpg)) + 
  geom_point() 
ggannotate(p)
```

